\section{The evidence-binding gap in AI reasoning systems}

Orchestrator-in-the-loop agentic systems route a user's request through a lead orchestrator that dispatches
tool-bearing sub-agents and narrates a result. Magentic-UI is a
representative instance: a lead orchestrator directs sub-agents and Model Context Protocol tools
with the human in the loop through co-planning, action guards, and memory~\cite{magenticui2025}.
This realizes, at agent scale, the mixed-initiative principle Horvitz set out, coupling
automated services with the user under explicit uncertainty about intent~\cite{horvitz1999}.
Treating the human and the orchestrator as communicative agents distinguished only by function, 
agency, interactivity, and influence, is the agent-agnostic stance human-machine communication
adopts for exactly this configuration~\cite{banks2020agentagnostic,walther2011cmc}, with agency and
control continually negotiated between human and system rather than fixed~\cite{gibbs2021agency}.
Magentic-UI is evaluated on task completion, interaction cost, and safety; it proposes no metric for
whether the orchestrator's narrated output is bound to the evidence its sub-agents actually gathered.
Because users treat such systems as social actors~\cite{gambino2020casa} and accept fluent narration
through heuristic rather than systematic processing~\cite{koban2023dual}, an unbound claim can pass
human review unchallenged, and calibrated reliance instead depends on signals the narration does not
expose~\cite{hoff2015trust}. That missing measurement is the evidence-binding gap, and it is the gap
this paper defines and measures.

Scientific and physical-domain outputs make claims requiring empirical grounding. On orca encounter
likelihood at Haro Strait, a language model may name the salmon corridor, acoustic monitoring
network, and behavioral studies, without binding any claim to generating observations, prior
analyses, or data artifacts. The response is fluent; it is not evidentially accountable. Fluency is not comprehension, and whether such systems understand what
they emit is itself contested~\cite{mitchell2023understanding}. Fluent interfaces routinely
manufacture an impression of accountability they have not earned, a banal deception embedded in
everyday human-machine contact~\cite{natale2021deceitful}, and critical scholarship holds that such
accountability gaps demand scrutiny beyond task-completion metrics~\cite{gardner2024critical}. An
output-level binding criterion sidesteps the comprehension dispute by measuring traceability rather
than adjudicating internal understanding.

Formally, the gap is the structural disconnection between stated claims and the observations,
methods, and experiments that would make them verifiable, not factual accuracy per se, but
whether a judge or downstream system can trace the claim to ground-truth data versus parametric
pattern reproduction.

Prior work on RAG hallucination reduction~\cite{arxiv230915217,arxiv230514627,arxiv240115884,arxiv231109533,arxiv230514292}
establishes that retrieval-augmented generation reduces factual error rates and that citation generation
can be automated. However, this literature focuses on whether RAG improves accuracy, not on whether
the improvement comes from evidence binding, from the system's outputs being traceable to their
generating context, versus improved parametric knowledge activation.

The geospatial grounding baseline makes this distinction concrete. Google Maps grounding achieves high
citation density for place-of-interest queries. For marine science evidence queries, the same domain,
but requiring scientific rather than place-level grounding, Maps grounding resolves 0 of 25 citations
to scientific or dataset sources. NOAA fisheries data, the Center for Whale Research census, Chinook
salmon migration corridors, and acoustic hydrophone recordings are all generated as claims but remain
entirely uncited. The system knows about these phenomena from training; it cannot bind its claims to
verifiable data artifacts.

$\Runcited$, the fraction of scientific sentences in a system's output with no bound
citation, is the measurement primitive that makes this gap quantitative, reproducible, and
architecture-diagnostic.
