\section{$\Runcited$ as a formal evaluation metric}

\subsection{Definition}

Let $S$ be the set of sentences in a system's output. Define $S_\mathrm{sci} \subseteq S$ as the
sentences containing scientific or quantitative markers. For each sentence $s \in S_\mathrm{sci}$,
let $\mathrm{cited}(s)$ be the indicator that $s$ names at least one entity appearing in the bound
citation set $C$.

\begin{equation}
\label{eq:runcited}
\Runcited = \frac{|\{ s \in S_\mathrm{sci} : \neg\,\mathrm{cited}(s) \}|}{|S_\mathrm{sci}|}
\end{equation}

$\Runcited = 0$ when every scientific sentence names a bound citation. $\Runcited = 1$ when no
scientific sentence does. It is undefined when $|S_\mathrm{sci}| = 0$.

\subsection{Properties}

$\Runcited$ has four properties that make it useful as an evaluation primitive.

\textbf{Claim-level granularity.} The metric operates at the sentence level. A response with 10
scientific sentences and 9 bound citations still has $\Runcited = 0.1$, distinguishing it from
response-level pass/fail metrics.

\textbf{Architecture-diagnostic.} Two systems with the same surface accuracy can have very different
$\Runcited$ values depending on whether their reasoning traces are passed as context to the language
model. The metric discriminates between evidence-bound and parametric-knowledge-based claim generation.

\textbf{Reproducible.} $\Runcited$ is computed from the raw API output. No human annotation is
required. The scientific marker set is implemented as a regex over domain-specific technical terms.

\textbf{Model-independent.} $\Runcited$ can be applied to any language model's output. It measures
the binding between output and context, not internal model behavior.

\subsection{Relation to prior evaluation frameworks}

The Ragas framework~\cite{arxiv230915217} evaluates faithfulness (whether claims are supported by
retrieved context), answer relevancy, and context precision. These metrics require a reference context.
$\Runcited$ measures a weaker but more deployable condition: whether the model's output names sources
in the context it was given, without requiring a predefined reference set.

The ALCE benchmark~\cite{arxiv230514627} evaluates citation generation at the claim level. $\Runcited$
is complementary: it measures the rate of claims carrying no citation at all, regardless of citation
validity.

$\Runcited$ is also distinct from explainable-AI evaluation, which concerns the transparency of a
model's decision process rather than the binding of its output to gathered evidence~\cite{clinciu2019survey},
and from reasoning-chain evaluation, which scores the internal correctness and informativeness of
intermediate steps rather than their traceability to artifacts~\cite{arxiv230410703}.
