\section{The four gaps in encounter forecasting}

The authoritative occurrence record for southern resident killer whales in the Salish Sea is built from opportunistic data. Olson et al. analyzed 82{,}447 sightings collected from 1976 to 2014 and resolved them into a small number of recurring hotspots, while stating that these opportunistic, citizen-sourced records yield reliable distribution estimates only ``if the potential for error and bias is taken into consideration''~\cite{olson2018}. That conditional is the load-bearing problem. Diggle, Menezes and Su established the statistical consequence directly: when the process that places observations is stochastically dependent on the process being mapped --- preferential sampling --- conventional intensity estimates are misleading unless the dependence is modeled explicitly, for example through a log-Gaussian Cox process~\cite{diggle2010}. Effort-corrected products for this population exist; the DFO advisory surface for the southern Salish Sea adjusts for observer effort and detectability and strengthens the result with passive acoustic detections~\cite{thornton2022}. These three results establish the data, the bias, and a correction. None carries the correction into an operational system that exposes, per displayed cell, the evidence that earned its confidence; four compounding failures remain in that gap.

The \textbf{epistemic gap} is that a map shows an intensity or probability without exposing what evidence earned it. Users of AI-assisted spatial forecasts systematically overtrust displayed confidence when the basis of that confidence is not made visible alongside the prediction~\cite{arxiv200102114}. Survey practitioners have documented that uncertainty is routinely suppressed in data-driven maps even when authors know it exists~\cite{arxiv190801697}. Maps are not inherently untrustworthy; the contract between evidence quality and displayed certainty is not explicit.

The \textbf{effort gap} is that sightings and acoustic detections conflate animal presence with observer presence. This is the preferential-sampling regime Diggle et al. formalize: shore observers concentrate near access points, ferries, and interpretable coastlines, so naive hotspot aggregation of the kind Olson et al. caution against learns where people watch, not where whales travel~\cite{diggle2010,olson2018}. Movement ecology models show that the classical mass-action approximation fails when observer search effort is spatially structured~\cite{arxiv190705902}, and the machine learning wildlife conservation literature treats effort correction as a central unresolved problem in field data pipelines~\cite{arxiv211012951}.

The \textbf{provenance gap} is that no existing operational system exposes a per-prediction drilldown from a displayed probability to the kernels, detections, and gate verdicts that produced it. A researcher cannot audit whether a displayed confidence reflects a fitted diel kernel that passed a phase-shuffled null test and a held-out deviance skill check, or a prior-dominated estimate with no model fit.

The \textbf{governance gap} is that citizen-science shore reports and automated model promotions enter operational forecasting narratives without explicit quarantine or signed human authority. The information quality literature on citizen science establishes that unmoderated crowdsourced data produces systematic quality failures when used directly in probabilistic models~\cite{arxiv190508289}. The Gravity Spy program at LIGO provides a concrete operational precedent: citizen classifier outputs are quarantined and validated before influencing detector characterization~\cite{arxiv161104596}.

Olson et al. supply the occurrence record, Diggle et al. supply the correction for its sampling bias, and the DFO advisory surface applies that correction for this population~\cite{olson2018,diggle2010,thornton2022}. None supplies an operational loop combining live forecast, gate battery, per-cell provenance, quarantined community ingest, immutable promotion audit, and primary transactional database. orcast occupies that residual.
