\section{A gate-bounded honesty model}

\subsection{Distributional family}

Encounter counts are sparse, overdispersed, and zero-inflated. The negative binomial distribution arises as a gamma-Poisson marginalization and provides two free parameters to capture overdispersion and random effects~\cite{arxiv11123605}. Hurdle and NB models are required for species abundance estimation when Gaussian assumptions fail on ecological count data~\cite{arxiv201016040}. Species distribution modeling frameworks identify structured noise and under-counting as reasons that count-specific distributional families are necessary~\cite{arxiv210208534}.

The kernel model uses the log-linear intensity, decomposing the log encounter rate into a baseline plus additive cyclic kernel contributions from diel, lunar, and seasonal covariates:

\eqintensity

See also the methodology of record.

\subsection{Null gating}

Each fitted cyclic kernel must beat a phase-shuffled null before contributing to displayed confidence. Phase shuffling permutes detection times across stimulus phases to generate the reference distribution. Point process models for spike trains~\cite{arxiv201004875} and the time-elapsed neural model~\cite{arxiv150602361} establish the mathematical foundation. \eqnulltest

Time-rescaling applies the compensator transformation; residuals should be exponential, and a KS test on them is the second gate component~\cite{arxiv160603988}:

\eqtrescaling

An open measurement gap (C-2): this methodology is documented in neuroscience and statistical literature but lacks an ecology-domain citation in the arXiv corpus. The whitepaper cites methodological precedents.

\subsection{CV skill and calibration}

Gate E5 requires positive held-out deviance skill

\eqdevskill

and gate E6 requires PIT calibration

\eqpit Negative skill means the model is less predictive than the null and displayed confidence must be withheld. The epidemiological calibration literature establishes that PIT-based recalibration reliably improves log score~\cite{arxiv211206305}. The abstention loss framework~\cite{arxiv210408236} formalizes the principle that a predictive system should withhold confidence when skill is low. Spatial and temporal CV must account for autocorrelation~\cite{arxiv200514263,arxiv190511744}.

\subsection{Effective confidence gate}

Displayed effective confidence is bounded by the product of gate outcomes:

\eqeffconf Failed gates reduce confidence and expose the failing integrity condition alongside the metric. The gate is not a visibility switch; the forecast is always shown.
